### 2016-08-01 supporting-information/supporting_info.tex
supp info, tables, edits

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Remove the twocolumn option to create a single column SI file if required. 
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnassupportinginfo}

\title{Supporting Information}
\author{Hough et al.}

\doi{genetics.XXXXXXXXXX}

\begin{document}
\onecolumn

\maketitle

\section*{Supporting Information (SI)}

This supporting information may also be accessed at gihub at https://github.com/houghjosh/XYdiversity

\subsection*{SI Tables}

\begin{table}[h!]
\centering
\caption{Population identities (ID) and location information for \textit{R. hastatulus} samples from Texas and North Carolina}
\begin{tabular}{lrrr}
Population ID & Location & Altitude & Latitude & Longitude \\
\midrule
TX-MTP & Mount Pleasant, Texas & 130	 & 33.17453 & 94.98799 \\
OK-RAT & Rattan, Oklahoma & 138 & 34.15755 & 95.41325 \\
TX-LIV & Livingston, Texas & 83 & 30.69947 & 94.79981 \\ 
LA-DER & De Ridder, Lousiana & 67 & 30.8941 & 93.3143 \\
TX-ATH & Athens, Texas & 145 & 32.18471 & 95.8032 \\
OK-WIL & Willis, Oklahoma & 211 & 33.89663 & 96.83533 \\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}

North Carolina				
Population ID	Location	Altitude	Latitude	Long
SC-PRO	Prosperity, South Carolina	126	34.10792	81.43711
GA-BEL	Belfast, Georgia	15	31.84293	81.28405
GA-STA	Statesboro, Georgia	78	32.45237	81.84849
SC-BRA	Branchville, South Carolina	34	33.25082	80.80761
FL-HAM	Hammock, Florida	3	29.06816	82.64664
GA-GLA	Gladys, Georgia	97	31.48198	83.23783

\subsection*{SI Figures}

\subsection*{SI Datasets}
links:

\end{document}