### 2016-08-02 Rnotebook/ne.Rmd
Rnotebook, major edits on Ne section

---
title: "Nucleotide diversity predictions"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

```{r}
Nf=100 #number of females
Nm=100 #number of males
N=Nf+Nm #total number of individuals
x=Nm/(Nf+Nm) #proportion of males

op <- par(mfrow = c(1, 2))
m=0
f=0

curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m))))/N, from=0, to=1, lty=2,
      xlab="proportion of males", ylab="NeA/N", main="Autosomes", ylim=c(0,1), xlim=c(0,1))

mh=0.5
curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+mh))))/N, from=0, to=1, add=T)

m1=1
curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m1))))/N, from=0, to=1, add=T)

m2=2
curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m2))))/N, from=0, to=1, add=T)

m3=3
curve(((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m3))))/N, from=0, to=1, add=T)

#for sex-linked genes
curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m))))/N, from=0, to=1, lty=2, xlab="proportion of males", ylab="Ne/N", main="Sex chromosomes", ylim=c(0,1),xlim=c(0,1))

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+mh))))/N, from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m1))))/N, from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m2))))/N, from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m3))))/N, from=0, to=1, add=T)

curve((((N*(x)*(1+f))/2))/N, from=0, to=1, col="red", add=T)

par(op)

```

```{r}

ol <- par(mfrow = c(1, 2))

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m))))/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m)))), from=0, to=1, lty=2,
      xlab="proportion of males", ylab="Relative Ne", main="X/A")

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+mh))))/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+mh)))), from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m1))))/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m1)))), from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m2))))/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m2)))), from=0, to=1, add=T)

curve(((9*N*(x)*N*(1-x))/((4*N*x*(1+f))+(2*N*(1-x)*(1+m3))))/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m3)))), from=0, to=1, add=T)

#unsure if this is right; substituting in the modified Nm's here based on crow model where Nm*(km-1))/((km-1)+(x/km)) is the effect of male rep variance on male Ne.

curve(((N*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m)))), from=0, to=1, col="red", lty=2, ylim=c(0,1), xlab="proportion of males", ylab="Relative Ne", main="Y/A")

curve(((80*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+mh)))), from=0, to=1, add=T, col="red")

curve(((66.66667*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m1)))), from=0, to=1, add=T, col="red")

curve(((50*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m2)))), from=0, to=1, add=T, col="red")

curve(((40*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m3)))), from=0, to=1, add=T, col="red")

par(ol)
```

```{r}
curve(((N*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m)))), from=0, to=1, col="red", lty=2, ylim=c(0,1), xlab="proportion of males", ylab="Relative Ne", main="Y/A")

curve(((80*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+mh)))), from=0, to=1, add=T, col="red")

curve(((66.66667*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m1)))), from=0, to=1, add=T, col="red")

curve(((50*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m2)))), from=0, to=1, add=T, col="red")

curve(((40*x*(1+f))/2)/((4*N*(1-x)*x*N)/((x*N*(1+f))+((1-x)*N*(1+m3)))), from=0, to=1, add=T, col="red")

```