### 2016-11-07 manuscript.tex
minor edit on last push

\documentclass[9pt,twocolumn,twoside]{gsajnl}
% Use the documentclass option 'lineno' to view line numbers
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{adjustbox}
\usepackage{geometry}
\articletype{gos} % article type

\title{Nucleotide Diversity Loss on a Plant Y Chromosome Following Recent Recombination Suppression}

\author[$\ast$,$\dagger$,1]{Josh Hough}
\author[$\dagger$]{Wei Wang}
\author[$\dagger$]{Spencer C.H. Barrett}
\author[$\dagger$]{Stephen I. Wright}

\affil[$\ast$]{Department of Plant Sciences, University of California, Davis}
\affil[$\dagger$]{Department of Ecology and Evolutionary Biology, University of Toronto}

%\affil[$\dagger$]{Author two affiliation}
%\affil[$\ddagger$]{Author three affiliation}
%\affil[$\S$]{Author four affiliation}
%\affil[$\ast\ast$]{Author five affiliation}

\keywords{Sex Chromosome Evolution; Nucleotide Diversity; Recombination; Deleterious Mutations}

\runningtitle{Plant sex chromosome evolution} % For use in the footer

\correspondingauthor{Josh Hough}

\begin{abstract}
X and Y chromosomes differ in effective population size ($N_{e}$), rates of recombination, and exposure to natural selection, all of which can affect levels of genetic variability within the genome. On Y chromosomes with suppressed recombination, selection is expected to remove linked neutral variation and reduce the $N_{e}$ of Y compared to X chromosomes or autosomes. However, non-selective factors including female biased sex ratios and high variance in male reproductive success can also reduce Y-linked $N_{e}$, making it difficult to infer the causes of low Y diversity. Here, we investigate the factors affecting X- and Y-linked neutral polymorphsm during plant sex chromosome evolution in \textit{Rumex hastatulus} (Polygonaceae), a dioecious annual with recently evolved sex chromosomes. Strikingly, we find that neutral diversity for genes on the Y is on average \textasciitilde 2.1\% of the value for their homologues on the X, corrsponding to a chromosome-wide reduction of \textasciitilde 93\% compared to the neutral expectation. We demonstrate that the magnitude of this diversity loss is inconsistent with a reduced male $N_{e}$ caused by neutral processes including female-biased sex ratios and high variance in male reproductive success. Rather, using forward population genetic simulations and estimates of the distribution of fitness effects of deleterious mutations, we show that Y diversity can be explained by purifying selection removing linked neutral variation, although selective sweeps may have also contributed. Given the recent origin of \textit{R. hastatulus} sex chromosomes, our study suggests that Y chromosome degeneration in the early stages occurs through strong selective interference among many linked sites rather than primarily through Y-inactivation followed by neutral genetic drift.
%abstract is already 270 words. Should be 250 (but might be ok). Just can't get longer.
\end{abstract}

\setboolean{displaycopyright}{true}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\marginmark
\firstpagefootnote
\correspondingauthoraffiliation{jhough@ucdavis.edu}
\vspace{-11pt}

\section*{Introduction}

\lettrine[lines=2]{\color{color2}M}{}orphologically distinct sex chromosomes have evolved multiple times independently in both plant and animal kingdoms \citep{westergaard1958,ohno1967,bull1983,charlesworth1991}. Despite clear biological differences between these kingdoms, X and Y chromosomes in both lineages have undergone similar genetic changes. For example, in both groups the loss of recombination between X and Y chromosomes has been associated with an accumulation of deleterious mutations and a gradual loss of genes from the Y chromosome \citep{hough2014,bergero2015,bachtrog2013NRG}, and in some species, this genetic degeneration of the Y has led to the evolution dosage compensation of the X \citep{charlesworth1996CB,muyle2012,mank2013sex,papadopulos2015}. The independent evolution of these phenomena across a broad range of distantly-related species suggest that general evolutionary mechanisms may be involved, but inferring the causes of molecular evolution and patterns of polymorphism in genomic regions that lack recombination is a longstanding challenge for both theoreticians and experimentalists \citep{charlesworth1978,feldman1980evolution,barton1995general,charlesworth1996CB,otto1997deleterious,charlesworth2000degeneration,mcvean2000effects}.
% and reduced genetic variation on Y chromosomes \citep{qiu2010nucleotide, ellegren2011,bachtrog2013NRG}.

One fundamental difference between the X and Y chromosomes is that there are 1/3 as many Y-linked gene copies as X-linked ones in a diploid population. Because of this, the effective population size ($N_{e}$) of the Y chromosome is expected to be 1/4 that of the autosomes, whereas the $N_{e}$ for the X chromosome should be 3/4 that of autosomes (assuming an equal number of reproducing females and males). The lowered $N_{e}$ of the Y chromosome implies that deleterious mutations are more likely to rise to high frequency because of gentic drift \citep{nei1981genetic} - an effect that can be exacerbated by selective interference among genetically linked sites (the "Hill-Robertson Effect"). These effects both predict a reduction in the level of neutral polymorphism maintained on the Y chromosome, which at equilibrium is proportional to the product of $N_{e}$ and the mutation rate, $\mu$ \citep{Kimura1984,charlesworth1987}.

In species with female-biased reproductive sex ratios or extensive male-male competiton, a high variance in male reproductive success can also reduce the $N_{e}$ of genes on the Y chromosome \citep{caballero1995,charlesworth2001,laporte2002,pool2007,ellegren2009}, suggesting that inferences about the effects of linked selection on the Y need to be distinguished from neutral models that include sex ratio variation as well as male-female differences in reproductive success. High variance in male reproductive success, for example, is predicted to reduce both Y-linked $N_{e}$ and autosomal $N_{e}$ \citep{kimura1964number,nomura2002effective}, and therefore evidence for this effect can be obtained by comparing levels of neutral variability on X and Y chromosomes relative to the autosomes. In particular, high variance in male reproductive success predicts a reduction in the Y/A diversity ratio, but an increase in the X/A ratio. Based on such comparisons, studies in humans have found examples of historical sex-differences in dispersal during human migration \citep{wilkins2006unraveling}, and the inflated X/A ratio has been suggested to reflect a historical excess of breeding females over males \citep{hammer2008sex}, though there is considerable variation in X/A estimates in human and other demographic scenarios have also been suggested \citep{bustamante2009,hammer2010}.

Despite widespread interest in determining the effects of selection patterns of sex-linked variability diversity \citep{ellegren2011,bachtrog2013NRG}, we still know very little about how linked selection has affected patterns nucleotide polymorphism. In humans and Drosophila \citep{mcallister1999,bachtrog2000}, where large population genomic data sets are readily available, studies examining Y chromosome variability have revealed that Y-linked diversity is considerably lower than predicted under models of neutral evolution \citep{hellborg2004,bachtrog2013NRG,Wilsonsayres2014}. However, it remains unclear whether such inferences about the effects of linked selection on the Y should also apply to systems in which sex chromosomes evolved \textit{do novo} from autosomes much more recently. The effects of diversity loss due to selection might be expeted to be lower in such systems because of a shorter history of recombination suppression, but contrary to this, simulations reveal that background selection and selective sweeps both have strongest effects during the early stages of Y-chromosome evolution, before Y-chromosomes have lost the majority of their genes \citep{bachtrog2008temporal}. Indeed, studies of Y-degeneration in the young plant sex chromosomes \textit{R. hastatulus} and \textit{Silene latifolia} have both revealved signs of rapid Y chromosome degeneration \citep{hough2014,papadopulos2015,charlesworth2016plant}. In these cases, if the early stages of Y-chromosome evolution involved widespread Y-linked gene-inactivation (silencing) \citep{orr1998adaptive}, then it is possible that the strength of linked selection would then be minimized such that gene loss was mostly driven by gene-silencing followed by neutral genetic drift \citep{bachtrog2013NRG} rather than selective interference due to widespread linkage disequilibrium.

%The strength of this effect is dependent on the rate of recombination between a selected locus and a neighbouring region; essentially, the effect becomes more pronounced with less recombination because of linkage disequilibrium between adjacent loci.

%On the other hand, if gene silencing and Y inactivation occur very early during the process of Y chromosome degeneration, it is possible that the strength of linked selection is minimized rapidly even on relatively young Y chromosomes, and that much degeneration may occur neutrally following the silencing of Y-linked genes (cite backtrog NRG review).

%because recombination suppression was relatively recent on plant Y chromosomes, it is an open question whether weak-HRI has had enough time to have a large effect. Although selective interference (indlucing BGS and Selective Sweeps) is expected to have the greatest effects during the early stages of Y-chromosome evolution \citep{bachtrog2008temporal} the R. hastatulus Y chromosome has undergone significant degeneration and gene loss over a short period of time, suggesting that it may behave more similarly to more ancient Y chromosomes.

%Background selection (purifying selection against deleter- ious alleles) [27] and selective sweeps (positive selection for beneficial variants) [28] both have the effect of reducing nucleotide diversity at linked sites.
%as modeled by \citep{kimura1964number} and \citep{nomura2002effective}.

%evolutionary theory predicts that neutral variability on Y chromosomes with suppressed recombination can be significantly reduced by Hill-Robertson Interference (HRI), which causes a reduction in $N_{e}$ of neutral sites that are genetically linked to sites experiencing selection. The major types of HRI include (i) selection on strongly beneficial mutations ("selective sweeps") \citep{smith1974hitch,aquadro1994} (ii) selection against strongly deleterious mutations) \citep{charlesworth1996background,charlesworth1994effect} \citep{charlesworth1996CB,charlesworth2000degeneration}, and (iii) intereference between weakly selected sites (weak selection Hill-Robertson interference). In all of these processes, linkage  disequilibrium between selected and neutral sites results in a level of neutral variation below the level predicted in the absence of selection.

%Consistent with an important role for Hill-Robertson interference in driving down diversity on the Y chromosome, s

%discussion point?:
%Moreover,

To investigate the factors affecting nucleotide diversity in the early stages of sex chromosome evolution we analyzed neutral polymorphism levels on X, Y, and autosomal chromosomes in the plant \textit{R. hastatulus }(Polygonaceae). This species is a dioecious annual with highly heteromorphic X and Y chromosomes that are estimated to have evolved  approximately 15 MYA \citep{quesada2011,grabowska2015,navajas2005}, making sex chromosomes in this species over 100 million years younger than the well-studied mammalian sex chromosomes \citep{lahn1999,ross2005dna}. \textit{Rumex hastatulus} has also received particular attention because of the unique occurrence of an intraspecific sex chromosome system, in which both XY and $XY_{1}XY_{2}$ males occur in geographically distinct 'chromosomal races' \citep{smith1963mechanism}. The $XY_{1}XY_{2}$ sex chromosome system in this species (the North Carolina race) is thought to have originated through an X-autosome fusion, with the XY system (the Texas race) maintaining the ancestral chromosome complement \citep{smith1964evolving}. Despite the recent origin of sex chromosomes in both races, a recent study revealed that both the ancestral and neo-Y chromosomes have undergone significant gene loss and functional deterioration \citep{hough2014}.  In order to simplify our comparison of polymorphism levels within the genome, we focus on the ancestral Y chromosome that occurs in both sex chromosome races.

Of particular relevance to the present study, \textit{R. hastatulus} populations have been found to consistently exhibit female-biased reproductive sex ratios, with a mean sex ratio of $N_{f}/(N_{m}+N_{f})=0.6$ \citep{pickup2013influence}. Female-biased sex ratios in dioecious plants with sex chromosomes are not uncommon (refs), and their occurrence in \textit{Rumex} provides an excellent opportunity to study both the demographic and selective effects contributing to sex-linked variability. Here, using genome-wide estimates of polymorphism together with forward population simulations, we test whether levels of neutral diversity on sex chromosomes and autosomes in this species are consistent with neutral evolutionary processes, including the joint effects of sex ratio bias and high variance in reproductive success, or whether the removal of linked neutral variation caused by selection on the Y has led to greater loss of diversity than expected under neutrality.

\section*{Materials and Methods}
\subsection*{Population Samples and Sex-Linked Genes}
We analyzed sex-linked and autosomal genes identified from Illumina RNA sequence data from 24 individuals (12 males and 12 females, with 1 male and 1 female from each of 12 populations; 6 populations of each sex chromosome race). Samples were collected in 2010 from throughout the native range of \textit{R. hastatulus} (locations in Table S1), and plants were grown in the glasshouse from seeds collected from open-pollinated females. We extracted RNA from leaf tissue using Spectrum Plant Total RNA kits (Sigma-Aldrich). The isolation of mRNA and cDNA synthesis was conducted according to standard Illumina RNAseq procedures, with sequencing conducted on two Illumina HiSeq lanes with 150-bp end reads at the Genome Quebec Innovation Center. Reads from these 24 samples were mapped to the \textit{R. hastatulus} reference transcriptome \citep{hough2014}, and raw sequences are available from the GenBank Short Read Archive under accession no. SRP041588. Reads were mapped using the Burrows–Wheeler Aligner \citep{li2010fast}, followed by Stampy \citep{lunter2011stampy}. We used Picard tools (http://picard.sourceforge.net) to modify mapping output into the format required for the Genome Analysis Toolkit \citep{mckenna2010genome} variant calling software, and subsequently removed genes with low coverage (<10x) and Phred Quality Scores <20. The population samples analyzed here were previously reported in Hough et al. (2014), where they were used to validate the ascertainment of sex-linked genes identified through segregation analysis. Here, to consider sex linked genes that were identified in each of our sequenced population samples, we focused on the set of 460 X/Y genes for which the Y homolog was found in both the Texas and North Carolina races (i.e., X/Y genes where the Y copy was inferred to be on the $Y_{1}$ chromosome).

\subsection*{Autosomal Genes}
In evaluating evidence for nucleotide diversity differences between X and Y chromosomes, it is important to distinguish between reduced Y-linked diversity, and the possibility that X-linked diversity is elevated above the level predicted from a neutral model. To do this, we normalized sex-linked diversity estimates by autosomal diversity, and compared empirical X/A and Y/A nucleotide diversity ratios to those predicted from neutral models and from simulations (described below). Because the criteria for identifying autosomal loci in Hough et al. (2014) were based on the occurrence of four segregating SNPs per locus, this set of genes is probably higher in diversity than the average autosomal gene. Therefore, for the present analysis we incorporated the larger set of all non-sex linked (putatively autosomal) genes from our transcriptome data. We also filtered this set to remove genes that may have been sex-linked but were not identified as such by Hough et al.'s conservative ascertainment criteria. In particular, we removed: (i) any genes in which there was evidence for at least one SNP with a sex-linked segregation pattern, (ii) any genes with SNPs showing fixed heterozygosity in males and fixed homozygosity in females, (iii) any genes with less than 10X coverage or greater than 100X coverage from independently obtained genomic coverage data (to filter out duplicate genes or those with highly repetitive sequences), and (iv) any genes containing SNPs with large (>0.4) allele frequency differences between males and females. Finally, we removed genes with fewer than 100 synonymous sites following this filtering to avoid biasing our results toward genes that may have been particularly short due to assembly problems. This filtering resulted in a set of 12,356 and 11,350 autosomal genes in the Texas and North Carolina races, respectively.

\subsection*{Phasing X and Y alleles}
To estimate polymorphism for the X and Y sequences separately, it is necessary to infer the phase of SNPs in sex-linked transcripts in males. In previous work, phasing alleles on \textit{R. hastatulus} sex chromosomes was achieved using segregation analysis from a genetic cross. Here, to phase SNPs from population samples where such segregation data was unavailable, we used HAPCUT \citep{bansal2008hapcut}, a maximum-cut based algorithm that reconstructs haplotypes using sequenced fragments (Illumina read data) from the two homologous chromosomes to output a list of phased haplotype blocks containing the SNP variants on each chromosome. Because the resulting haplotype blocks produced by HAPCUT contained SNPs that were phased relative to each other, but not designated to either the X or Y chromosome, we assigned individual variants to X or Y by independently identifying fixed X-Y differences with each haplotype block (i.e., sites in which all females were homozygous, and all males were heterozygous). Identifying such fixed differences within phased haplotype blocks enabled us to then infer the correct phase (X or Y) of the polymorphisms from HAPCUT’s output. This was done by matching the phase of fixed X-Y differences with their neighboring polymorphic sites: when a fixed X-Y difference occurred in the same phased haplotype block as a polymorphic site, the polymorphic variants in that block were assigned to either X or Y based on the known phase of the fixed difference with which they were matched. SNPs that were identified outside of phased blocks, or in blocks without fixed X-Y differences, were recorded as missing data. Finally, we filtered out SNPs with coverage > 60, QUAL score > 60, and those within a distance of 10bp or less from indels. This procedure was conducted using a combination of Perl and Bash scripts, and resulted in fasta-formatted alignments of X and Y sequences for 372 sex-linked genes from the 24 individuals in our study.

We further validated the results of HAPCUT’s allele phasing by comparing the accuracy of this method with the phasing-by-segregation method that was conducted in Hough et al. (2014). To do this, we first phased the sequence data from parents and their progeny using HAPCUT’s algorithm (using the same parameters as for the population data), and then identified cases where SNPs were inferred on the Y chromosome by HAPCUT, but where the ‘true’ level polymorphism was known to be zero. We identified 7 percent of sex-linked genes with phasing errors of this kind, or which were otherwise determined to have genotyping errors resulting in false SNP calls. This corresponds to a SNP error rate estimate of 1.7 x 10-4. Note that this rate is very low relative to population-based estimates of polymorphism on the X and autosomes (Table 1), and therefore should have minimal effects on our estimation of X/A polymorphism. However, because the rate is high relative to the expected level of true polymorphism on the Y-chromosome, we  further filtered genes in which we found evidence for false-positive Y-polymorphisms arising from: (i) phasing errors caused by gene duplicates (more than two haplotypes), (ii) polymorphisms around indels, and (iii) genotyping errors caused by low Y-expression. This filtering was done by manually inspecting sequences in IGV \citep{robinson2011integrative} and identifying each individual putative polymorphism on the Y chromosome.

\subsection*{Estimating nucleotide diversity on sex chromosomes and autosomes}
For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}. To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (\X Figure 2; Table 1). We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping per gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each. Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race. Note that the lack of recombination on the Y chromosome implies that assumptions about independence across loci are violated, suggesting that the true uncertainty in our Y/A diversity estimate may be wider than implied by bootstrapping across genes. To address this issue, we also used a maximum likelihood approach, implemented in a modified version of the MLHKA software \citep{wright2004hka} to independently estimate a credibility interval for the Y/A ratio using coalescent theory (Figure S1). Because of the thousands of genes involved, a likelihood method incorporating divergence to control for heterogeneity in mutation rate is not feasible, since this involves maximizing the likelihood estimate of the mutation rate for each locus independently.  Therefore, we assumed no heterogeneity in mutation rate, no recombination between Y-linked genes, and free recombination between autosomal loci. Our model therefore had two parameters; $\theta_{autosomal}$ and the Y/A ratio. We varied both parameters across a grid of points and evaluated the likelihood, where the Y/A ratio varied from 0.001 to 1, and $\theta_{autosomal}$ varied from 0.001 to 0.01.
%We also tested whether estimates of diversity for the X chromosome calculated from phased sequences from females were consistent with estimates from phased sequences from males. As no significant difference was observed (\X Figure? Table?), we report only results from females.

%Finally, we tested for population substructure within and between the two sex chromosome races to control for the possibility that hidden substructure in our sampled populations could affect estimates of diversity. To do this, we constructed neighbor-joining trees of sex-linked sequences from all populations in our study. The analysis was conducted on an alignment of the X- and Y-linked genes from \textit{R. hastatulus}, with orthologous autosomal sequences from the non-dioecious but closely related outgroup species \textit{R. bucephalophorus} used to root the tree (Figure S2). We used the Neighbor-Joining method \citep{saitou1987neighbor}, with evolutionary distances computed using Maximum Composite Likelihood \citep{tamura2011mega5}. The inferred trees revealed strong support for Y-linked genes from the $XY_{1}XY_{2}$ ("North Carolina") race being paraphyletic (98\% bootstrap support), with samples from two populations (hereafter the "SC sub-clade") forming a monophyletic group, and samples from Florida and Georgia (hereafter the "FL sub-clade") also forming a monophyletic group, that was more closely related to the XY ("Texas") race (Figure S2). In light of this evidence for significant population substructure, and to consider between-population differences, we estimated X and Y diversity in each of the three sub-clades (SC, FL, TX) separately. For model-based analyses, we focused on the population sample from the Texas XY race, given that this reflects the presumed ancestral karyotype and has the largest sample size.

\subsection*{Neutral predictions and the effect of sex ratio bias on diversity}
To test whether the levels of diversity we observed on X, Y and autosomal chromosomes could be explained by the occurrence of biased reproductive sex ratios or high variance in reproductive success, we compared our empirical diversity estimates to predictions from a neutral model that jointly considers these two effects.

In particular, the effective population size for genes on each chromosome can be given by  \citep{wright1931evolution}:

\begin{equation}
N_{e{A}} = \frac{4N_{m}N_{f}}{N_{m}+N_{f}}\label{eq:NeA}
\end{equation}

\begin{equation}
N_{e{X}} = \frac{9N_{m}N_{f}}{4N_{m}+2N_{f}}\label{eq:NeX}
\end{equation}

\begin{equation}
N_{e{Y}} = \frac{N_{m}}{2}\label{eq:NeY}
\end{equation}

 Using these predictions, we calculated the expected X/A and Y/A ratios of diversity for the estimated \textit{R. hastatulus} sex ratio (N_{f}/$N_{m}+N_{f})=0.6$ \citep{pickup2013influence}, assuming that the level of neutral polymorphism is given by the product of the mutation rate and the effective population size, $\theta=4N_{e}\mu$ \citep{watterson1975}. Sex-specific variance in reproductive success can lead to departures from this prediction, and so we also evaluated whether any ratio of male- to female- effective population sizes could explain our observed combination of X/A and Y/A diversity.

\begin{figure}[htbp]
\centering
\includegraphics[width=\linewidth]{Figure1.png}
\caption{The relation between relative effective population size and sex ratio bias for genes on autosomes (\textbf{A}) and sex chromosomes (\textbf{B}) (Y chromosome in red). The sex ratio is shown as the proportion of males, $N_{m}/(N_{f}+N_{m})$, where $N_{m}$ and $N_{f}$ are the effective number of breeding males and females, respectively, plotted against $N_{e}/N$, where $N=N_{m}+N_{f}$ and the $N_{e}$ for sex chromosomes and autosomes is given in equations 1-3. Dotted curves show predictions in the standard neutral model, where both males and females produce Poisson-distributed offspring numbers and the chromosomal $N_{e}$'s are given by equations 4-6 \citep{wright1931evolution}. Solid curves correspond to increasing levels of variance in male reproductive success \citep{nomura2002effective} (see Methods). Assuming $\theta=4N_{e}\mu$ and equal neutral mutation rates among genes, the predicted $N_{e}$'s are used to generate null predictions for X/A and Y/A ratios of diversity.
}
\label{fig:spectrum}
\end{figure}

\subsection*{Simulations of purifying selection}
To test whether our observed level of Y-chromosome diversity could be explained by the effects of linked purifying selection, we followed the approach used in \citep{Wilsonsayres2014} and conducted forward-time simulations of a nonrecombining Y chromosome using the software SFSCODE \citep{hernandez2008flexible} and compared our empirical diversity estimates on the Y with those from simulations. This approach was implemented rather than using analytical predictions of background selection primarily because the equilibrium background selection model over-predicts the reduction in diversity when there are many linked sites under selection \citep{KaiserCharlesworth}, as is expected to be the case for large Y chromosomes that lack crossing over.

To parameterize the forward simulations, we estimated the distribution of fitness effects of deleterious mutations (DFE) using our polymorphism data on the X-chromosome with the program DFE-alpha (Keightley and Eyre-Walker ref). This program fits a gamma distribution of selection coefficients to the observed frequency distribution of nonsynonymous and synonymous polymorphism from the population genetic analyses above. We used this estimated gamma distribution in our forward simulations. To examine the expected reduction in diversity, we varied the number of sites under selection, and obtained an approximate likelihood surface for our observed data based on the proportion of simulations that matched the observed estimate of $\pi_{synonymous}$ . To make the simulation output comparable to our data, we initialized the simulations with our empirically estimated autosomal $\theta$, adjusted for the expected neutral reduction in effective population size on the Y chromosome under a sex ratio of Nf/(Nf+Nm)=0.6, and we sampled 6 haploid chromosomes per simulation. Simulated sequences contained 45,331 of linked neutral sequence from which we calculated diversity, matching the number of synonymous sites sampled in our empirical analyses on the Y chromosome (see Supporting Information).

\section*{Results and Discussion}

\subsection*{Extensive loss of Y-chromosome diversity}
Our analysis reveals that diversity on the \textit{R. hastatulus} Y chromosome is significantly lower than expected under neutrality, with estimates indicating Y/A=0.02, ....which is 12.5 fold lower than the standard neutral prediction of Y/A = 0.25 (P<0.0001). We also observe that the Y chromosome shows a 40-fold lower than mean diversity on the X chromosome (Table 1). Note that by normalizing X and Y diversity by autosomal diversity, our results indicate that the X-Y difference we observed was not due to an elevation of X chromosome diversity, but rather a Y-specific reduction. Conceivably, such low diversity on the Y could arise from a low Y-linked mutation rate, or a lower mutation rate in males compared to females. However, these possibilities are unlikely because the number of synonymous mutations in X and Y lineages, estimated by both parsimony and maximum likelihood, are not significantly different \citep{hough2014}. Similarly, our estimates of weighted average synonymous substitution rate between \textit{R. hastatulus} and the outgroup \textit{R. bucephalophorus} reveal similar levels of synonymous divergence at sex-linked (0.2016) and autosomal genes (0.219) \citep{hough2014}.

%After initial analyses revealed a much higher level of diversity on the Y chromosome of the North Carolina race (SUP FIG), we examined evidence for population structure using a Neighbor-joining tree (sup fig). Surprisingly, this analysis revealed that the Y chromosome from this race appears to be paraphyletic, with XYY samples from Florida being more closely related to the TX race than either is to samples from South Carolina (suppfig). Once we examine diversity for these two subclades separately, a strong loss of diversity on the Y chromosome is once again apparent (Table 1).

%Although our sampling of each \textit{R. hastatulus} sub-clade is limited, the discovery of three phylogenetically distinct monophyletic groups is interesting because it suggests the possibility that introgression occurred between the ancestral Texas (XY race) and the derived North Carolina ($XY_{1}Y_{2}$) race, leading to a secondarily-derived $XY_{1}Y_{2}$ sub-clade. As the North Carolina and Texas races are known to be inter-fertile \citep{smith1964evolving}, we suggest that the Florida sub-clade inferred here likely originated through hybridization between a female from the SC clade harboring the X-autosome fusion, and a male from the XY Texas race.

\begin{table}[t!]
\centering
\caption{Estimates of neutral diversity by race on \textit{R. hastatulus} sex chromosomes and autosomes.}
\begin{tabular}{ccccccccc}
\textbf{} & \multicolumn{2}{c}{\textbf{Texas}} & \multicolumn{2}{c}{\textbf{South Carolina}} & \multicolumn{2}{c}{\textbf{Florida}} \\
chromosome & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ & $\theta$ & $\theta/\theta_{A}$ \\
\midrule
A & 0.006 & 1 & 0.006 & 1 & 0.005 & 1 \\
X & 0.0047 & 0.85 & 0.0019 & 0.33 & 0.0018 & 0.37 \\
Y & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 & $\num{e-4}$ & 0.002 \\
\addlinespace

\bottomrule
\end{tabular}
\end{table}

%Our results also indicate a significant reduction in X/A diversity in the derived SC and FL sub-clades of the North Carolina race ($X/A_{FL}=0.33$ and $X/A_{SC}=0.37$) compared to the Texas race ($X/A_{TX}=0.85$) (Figure 2). Although not expected, this reduction in diversity may be associated with the recent origin of the $XY_{1}Y_{2}$ sex chromosome system, which is thought to have originated through an X-autosome fusion involving the ancestral 3rd chromosome in the Texas race \citep{smith1964evolving}. Evidence supporting this autosomal origin was recently obtained by \citep{grabowska2015}, who reported that the ancestral third chromosome in the Texas race carries the 5S rDNA locus, which is now found on both the neo-X and the $Y_{2}$ sex chromosomes in the derived North Carolina race. If recent positive selection was involved in driving the evolution of this X-A fusion, which theory suggests can drive the evolution of such fusions \citep{charlesworth1980sex}, then the formerly autosomal segment on the X chromosome in the $XY_{1}Y_{2}$ sub-clades may have experienced a strong selective sweep, resulting in reduced X-linked diversity in the derived $XY_{1}Y_{2}$ sub-clades. However, this would require extensive recombination suppression between the fused and unfused X chromosomes, and it is also possible that sex-specific demographic history has driven these patterns. It will be important for future work to investigate in more detail the factors driving the establishment of the X-autosome fusion in this species, and how they might impact patterns of X-linked neutral diversity.

In contrast to the X chromosome, however, our data indicate a strong and consistent diversity reduction on the Y chromosome: an approximately 50 fold reduction compared to the mean $\theta_{Aut}$. We next consider several possible models - neutral and selective - that might explain this reduction.

\subsection*{Female biased sex ratio and high variance in male fitness}

The occurrence of female-biased sex ratios in \textit{R. hastatulus} has been predicted to lower Y diversity due to its effects on reducing male $N_{e}$ and therefore the neutrally expected $N_{e}$ of the Y chromosome (\X ref). This reduction $N_{e}$ on the Y is expected to be further accenuated with high variance in male reproductive success (Figure 1), which is not unusual in annual wind-pollinated plants such as \textit{R. hastatulus} that exhibit extensive phenotypic plasticity in plant size and flower production (\X Harper 1977). Moreover, given that male plants in this species produce large amounts of pollen, and that female flowers are uniovulate, we expected strong competition among males to fertilize females.
%Evidence for sex di¡erences in the variance of reproductive success can therefore be obtained by measuring levels of silent variability; strong sexual selection on males should bring the X-linked value close to, or even higher than, the auto- somal value (Charlesworth 1996b).

In common with most flowering plants we do not have marker-based estimates of the variance in male reproductive success in \textit{R. hastatulus}. However, by comparing our empirical estimates of diversity to predictions from models that jointly predict the effects on diversity of sex ratio bias and male reproductive variance, we evaluated whether these effects could jointly explain the level of Y/A diversity that we observed (see Methods). Conditioning on estimates of sex ratio bias in \textit{R. hastatulus}, ranging from $N_{m}/(N_{m}+N_{f})=0.4$ to $N_{m}/(N_{m}+N_{f})=0.35$ \citep{pickup2013influence}, the predicted Y/A diversity ratio  is approximately 0.2 (Table1, Figure 3). This is significantly lower than our estimated mean Y/A ratio of 0.02 ($\textit{P}<0.0001$), and remains significant even if we consider an upper bound estimate obtained from our likelihood based estimated of the confidence interval. This suggests that the sex ratio effect alone is insufficient to explain our data under the standard sex ratio model.

Assuming that there is extensive variance in male reproductive success (\X), the predicted ratios of Y/A diversity (\X supp) were also significantly higher than our estimates for the empirically estimated sex ratio (\X table) . We also find, however, that purely neutral models in which the sex ratio was highly female-biased (\X), with level of variance in male reproductive success on the order of (\X), predicted a Y/A ratio that could not be rejected (\X). However, because a highly females biased sex ratio is expected to increase the X/A ratio as well, these models simultaneously predicted a range of X/A ratios that were significantly different from what we observed (\X Figure). Thus, our results indicate that the combined effects of sex ratio bias and variance in reproductive success cannot jointly explain our observed levels of X, Y, and autosomal diversity.

\subsection*{Background Selection and Selective Sweeps}

%Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation.

%This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].

%hat genetic variability in these regions can be explained by interference among strongly deleterious mutations

%less effective in influencing the behaviour of neighbouring sites as the number of closely linked sites on a chromosome increases.

%For this chromosome, BGS predicts values of the ratio of neutral diversity to the rest of the genome of  0.1% [9], whereas the observed mean value is  6.5%

%The agreement is reasonably good given the uncertainties involved.

%silent site diversity for 22 genes on the D. miranda neo-Y is
%is  1.2% of the value for their homologues on the recombin- ing neo-X chromosome

% Our observed levels of diversity are essentially at that saturation point, highlighting that we are indeed in a realm where there are a large number of sites subject to purifying selection. So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome
%This reduces their Ne, so that they behave as if they were subject to weaker selection than with free recombination, thereby reducing their effects on linked neutral or nearly neutral variants [9].

%hat genetic variability in these regions can be explained by interference among strongly deleterious mutations

%less effective in influencing the behaviour of neighbouring sites as the number of closely linked sites on a chromosome increases.

%For this chromosome, BGS predicts values of the ratio of neutral diversity to the rest of the genome of  0.1% [9], whereas the observed mean value is  6.5%

%The agreement is reasonably good given the uncertainties involved.

%silent site diversity for 22 genes on the D. miranda neo-Y is
%is  1.2% of the value for their homologues on the recombin- ing neo-X chromosome

% Our observed levels of diversity are essentially at that saturation point, highlighting that we are indeed in a realm where there are a large number of sites subject to purifying selection. So I think we can safely say that, with the caveat about parameter uncertainty, the number of sites under selection is likely to be 800kb or more, consistent with ongoing purifying selection driving down diversity on this chromosome

%-Although we do not exclude the possibility that positive selection has also reduced Y chromosome variability, our simulations suggest that the reduction in diversity arising from purifying selection is sufficient to explain our observed patters of neutral variation. Given the recent origin of \textit{R. hastatulus} sex chromosomes, our results suggest that the low level of sequence diversity commonly observed on ancient Y chromosomes might evolve relatively quickly after sex chromosomes originate, with selection against deleterious mutations playing an important role during the earliest stages of their evolution.\end{abstract}

%A predicted consequence of a reduction in Ne is an accelerated rate of ¢xation of slightly deleterious substitu- tions, and a reduction in the rate of ¢xation of slightly advantageous mutations (Kimura 1983). Thus we might expect an increased rate of amino-acid replacement rela- tive to silent substitutions on evolving Y chromosomes (provided that su¤cient time has elapsed since the origin of the sex-chromosome system), and more ¢xations of non-preferred versus preferred synonymous substitutions at silent sites in exons (Charlesworth 1996a).

%It is worth noting that our results only apply to the $Y_{1}$ chromosome, as estimates of diversity were calculated for sex-linked genes that were shared between the $XX/XY$ and the $XX/XY_{1}Y_{2}$ sex chromosome systems. Previous work found that both the $Y_{1}$ and the more recently evolved "neo" $Y_{2}$ chromosomes exhibited signs of genetic degeneration, including gene loss, loss of expression, and an accumulation of amino acid-changing mutations \citep{hough2014}, but we cannot say from the present study whether the neo-Y chromosome has also undergone a reduction in diversity. However, the extensive reduction in diversity estimated on the $Y_{1}$ chromosome occurs in each of the three \textit{R. hastatulus} sub-clades (\X Figure 2; Table 1; Figure S2), suggesting that this effect is not population-specific.

% A further interesting aspect of the results from the S. latifolia Y chromosome is that they do not readily ¢t a recent selective sweep. Such an event would eliminate all diversity, so that existing SLY1 polymorphisms must have accumulated since the selective sweep and should thus mostly be at low frequencies (see table 1). However, all six SLY1 polymorphic sites found are non-singletons, and no signi¢cant deviations from neutrality are detectable by standard test statistics (Fu & Li 1993). However, this does not take into account the probable population subdivision in this species. Although a selective sweep eliminating variability throughout the worldwide population of this species seems unlikely, e¡ects of selective sweeps within local populations cannot be excluded.

%because the lack of recombination means that all genes on the Yof any given species share a common genealogy, which limits the amount of information that can be obtained from data on within-species variability.

%This raises the further issue of whether Y-chromosome degeneration is driven mainly by the accumulation of deleterious mutations on the Y or is due to processes turning down the expression of genes on an evolving Y.

%We have focused on the ¢rst of these possibilities, but it might well not completely explain Y degeneration. If mutations of relatively minor ¢tness e¡ects, with Nes not much bigger than 1, become ¢xed more rapidly on a degenerating Ychromosome than on the X, selection pres- sure would develop to reduce the transcription of Y genes, provided that this was compensated for in the hetero- gametic sex by an increased activity of their counterparts on the X.

\section*{Conclusions}

The observation of widespread degeneration and diversity loss on Y chromosomes illustrates the importance of recombination for maintaining fitness and the genetic variability needed for adaptation (Maynard Smith 1978; Kondrashov 19 93; Barton & Charlesworth 19 9 8). The extensive loss of diversity on a young plant Y chromosome revealed by our study provides a clear example of the .... a large genomic regions that lack recombination.

\section*{Acknowledgments}

\bibliography{bibliography}
\end{document}