### 2016-11-09 temp/fwdpy_sims/pwdpp2.Rmd

---
output:
  html_notebook: default
  pdf_document: default
---

###Can background selection explain low Y diversity in _R. hastatulus_?

We have done similations of background selection on the Y chromosome to test whether our observed level of Y/A diversity, $\theta_Y$/$\theta_A$, is consistent with the effects of background selection on this chromosome.

####Let's ask: given our simulation parameters, what does the theory predict?

Here are the input parameters:

* $Ne_Y = 1/4 Ne_A$, where the inital $Ne_A = 1000$
* $\mu = 0.005$ (neutral mutation rate, pre gamete, per generation)
* $U = 0.02$ (deleterious mutation rate, per _diploid_; so, twice the per gamete rate)
* $r = 0$ (recombination rate)
* $h = 1$ (dominance)

Under the BGS model, the expected $\pi$ is $E[\pi]=\pi_0e^{-\frac{U}{2sh+r}},$ $U$ is the mutation rate to strongly-deleterious alleles, $\pi_0$ is the value expected in the absence of BGS (_i.e._ $\pi_0 = \theta = 4N_e\mu$), $s$ and $h$ are the selection and dominance coefficients, and $r$ is the recombination rate.

For our parameters, we have:
$\pi_0 = \theta = 4N_e\mu = 5$, so

$$E[\pi] = 5e^{-\frac{0.02}{0.1}} = 1.839397$$

```{r}
5*(exp(-(0.02/((2*0.01)))))
```

Analagously for the autosomes, with $r = 0.5$, we have

$$\pi_0 = \theta = 4N_e\mu = 20$$

and

$$E[\pi] = 20e^{-\frac{0.02}{0.1+0.5}} = 19.24537$$

```{r}
20*(exp(-(0.02/((2*0.01)+0.5))))
```

and for the X chromosome

$$\pi_0 = \theta = 4N_e\mu = 15$$

and

$$E[\pi] = 15e^{-\frac{0.02}{0.1+0.5}} = 14.43403$$

```{r}
15*(exp(-(0.02/((2*0.01)+0.5))))
```

#### Expected Y/A and X/A diversity

OK, so then the expected ratios of diversity under background selection are:

$$\frac{E[\pi]_Y}{E[\pi]_A} = \frac{1.839397}{19.24537} = 0.095$$

and

$$\frac{E[\pi]_X}{E[\pi]_A} = \frac{14.43403}{19.24537} = 0.75$$

#### Simulated Y/A and X/A diversity

How does this compare with the simulation output?

import the simulation data

```{r}
#autosomes
a <- read.csv("/Users/josh/Dropbox/temp/Ydiversity/simdata/a_fixed.csv", header = T)
#X-chromosome
x <- read.csv("/Users/josh/Dropbox/temp/Ydiversity/simdata/x_fixed.csv", header = T)
#Y-chromsoome
y <- read.csv("/Users/josh/Dropbox/temp/Ydiversity/simdata/y_fixed.csv", header = T)
```

Let's see some plots of the simulated $\pi_Y$, $\pi_X$, and $\pi_X$

```{r}
boxplot(a$thetapi, x$thetapi, y$thetapi, names = c("A","X","Y"),
        ylab = expression(paste("Nucleotide Diversity (", pi, " ) ")))
```

The following $\pi$ values are the means from all of the simulations, for each chromosome type. You can see for the autosomes and the X chromosome, they are quite close to the theoretically expectd  values calculated above.

$$\pi_{A_{sim}} = 19.5$$
$$\pi_{X_{sim}} = 14.9$$
$$\pi_{Y_{sim}} = 4.2$$

```{r}
mean(a$thetapi)
mean(x$thetapi)
mean(y$thetapi)
```
However, the estimate for the Y chromosome is much higher than predicted the by model.

Moreover, ratios for the simulated data:

$$ \large X/A = \frac{\pi_{X_{sim}}}{\pi_{A_{sim}}} = 0.76$$

and

$$ \large Y/A = \frac{\pi_{Y_{sim}}}{\pi_{A_{sim}}} = 0.21$$

are not different from neutral expectations

```{r}
mean(x$thetapi)/mean(a$thetapi)
mean(y$thetapi)/mean(a$thetapi)
```

This is basically unchanged (well, even higher for the Y) if one weights these means by the number of polymorphic sites (called "npoly" in the data)

---
_Stephen: I don't know why this is? It might be some issue with the simulation somehow still treating sites on the Y as independent (this is my suspicion), but I have been investigating all day, and not quite sure anymore..._